JD notes on effects

### 2022-07-08 manuscript/body.tex
Trying to work on plane

@@

\item \textbf{Model variables}: Refer to the X-variables which go into the regression model; and represent columns in the model matrix. Each input variable may correspond to one or more model variable. In particular, variables with more than two categories (for instance religion in our previous example), or input variables modeled with interactions, a spline or polynomial interactions, will correspond to more than one model variable. For example, the regression models described by \EREF{simple_inter_higher_no_interaction} and \EREF{simple_inter_higher} have $4$ model variables -- $x_1, x_2, x_3$ plus intercept term and $5$ model variables -- $x_1, x_2, x_3, x_2x_3$ plus the intercept term, respectively.

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in \FREF{justify_plots}B. In simple linear models with no interactions, the mean of the input variable can be used as a proxy for the center point \jd{Is it a “proxy” or is it the thing??}. If there are more than one model variables associated with the focal input variable, however, the model center does not correspond to any given value of the focal predictor \jd{This seems like bad repetition}, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see \FREF{pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal predictors, when estimating the predictions and effects. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions (for bias correction) especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect confidence intervals. The anchor choice does not affect the central estimates. Typically chosen as the center point of the model variables corresponding to the focal predictor but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the predictions. We can choose any anchor, but the center point is a natural and stable choice.
\item \textbf{Marginal effects:} In simple linear models with no interaction terms, the coefficient estimates are simple and directly interpretable as the expected change in outcome for a unit change in focal input variable. This is the \emph{unconditional marginal effect} and it is constant across all the observations and levels of all other input variables. Consider models described by \EREF{simple_inter_higher_no_interaction} and \EREF{simple_inter_higher}. The marginal effect of $x_2$ in model~\ref{eq:simple_inter_higher_no_interaction} is $\frac{\partial y}{\partial x_2} = \beta_2$. On the other hand, the marginal effect of $x_2$ in model~\ref{eq:simple_inter_higher} is given by $\frac{\partial y}{\partial x_2} = \beta_2 + \beta_{23}x_3$. In other words, if there are no interactions, the marginal effect of $x_2$ on $y$ is constant, while, if there are interactions in the model, the marginal effect of a change in $x_2$ on $y$ depends on the value of the other \emph{conditioning} predictor, $x_3$. In essence, computation of marginal effects requires the use of partial derivatives and can be considered as the slope of a multi-dimensional surface with respect to the surface of interest \cite{leeper2017interpreting}. 
\end{itemize}

@@

To illustrate mean-based approach, consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function, where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the central estimate with respect to the focal predictor in question \citep{fox2009effect}; where $\bX^\star$ is the centered model matrix constructed by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.

>>>>

\jd{I guess this list should really be a desc, not an itemize.}

@@

\item \textbf{Model variables}: Refer to the variables which are combined to make the linear predictor (corresponding to the columns in the model matrix). Each input variable may correspond to one or more model variable. Input variables with more than two categories (for instance religion in our previous example), or input variables with non-linear response functions (e.g., spline or polynomial) will correspond to more than one model variable -- we call such variables “multi-parameter variables”.

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in \FREF{justify_plots}B. In simple linear models with no interactions, the mean of the input variable can be used as a proxy for the center point \jd{Is it a “proxy” or is it the thing??}. If there are more than one model variables associated with the focal input variable, however, the model center does not correspond to any given value of the focal predictor \jd{This seems like bad repetition}, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see \FREF{pred_cubic_plots}.

\item \textbf{Reference point:} The values (or sets of values) chosen for non-focal predictors, when estimating the predictions and effects. Typically the center point, but can also be chosen as a baseline value or as a mean across categories. We will also discuss using a set of quantiles or observations as a reference.

\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect-style confidence intervals. The anchor choice does not affect the central estimates, nor prediction-style. Often chosen as the center point of the model variables corresponding to the focal predictor.

\item \textbf{Marginal effects:} \jd{This isn't really one of our definitions; maybe some of this belongs in the text part.} In simple linear models with no interaction terms, the coefficient estimates are simple and directly interpretable as the expected change in outcome for a unit change in focal input variable. This is the \emph{unconditional marginal effect} and it is constant across all the observations and levels of all other input variables. Consider models described by \EREF{simple_inter_higher_no_interaction} and \EREF{simple_inter_higher}. The marginal effect of $x_2$ in model~\ref{eq:simple_inter_higher_no_interaction} is $\frac{\partial y}{\partial x_2} = \beta_2$. On the other hand, the marginal effect of $x_2$ in model~\ref{eq:simple_inter_higher} is given by $\frac{\partial y}{\partial x_2} = \beta_2 + \beta_{23}x_3$. In other words, if there are no interactions, the marginal effect of $x_2$ on $y$ is constant, while, if there are interactions in the model, the marginal effect of a change in $x_2$ on $y$ depends on the value of the other \emph{conditioning} predictor, $x_3$. In essence, computation of marginal effects requires the use of partial derivatives and can be considered as the slope of a multi-dimensional surface with respect to the surface of interest \cite{leeper2017interpreting}.

\item \jd{Should we consider defining effects-style and prediction-style CIs? Can we think of better words?}

\end{itemize}

@@

\jd{This should be more general. The natural way to do the comparison is to first write out the two ways of averaging, and then point out that they're the same when $g$ is the identity. In other words, do the general derivation before making the assumption. It doesn't ever seem good to write $g$ and then \emph{define} it as the identity.}
To illustrate mean-based approach, consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function, where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the central estimate with respect to the focal predictor in question \citep{fox2009effect}; where $\bX^\star$ is the centered model matrix constructed by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.
